---
title: "Assignment 2"
author: "Amruta"
date: "3/10/2020"
output:
  html_document: default
  word_document: default
---

The dataset is imported from IMDB having Positive and Negative feedbacks. This reviews are differenciated as positive and negative.

```{r}

library(keras)

imdb_dir <- "C:/Users/12342/Documents/AdvancedML/Assignment/aclImdb"
train_dir <- file.path(imdb_dir, "train")
labels <- c()
texts <- c()
for (label_type in c("neg", "pos")) {
  label <- switch(label_type, neg = 0, pos = 1)
  dir_name <- file.path(train_dir, label_type)
  for (fname in list.files(dir_name, pattern = glob2rx("*.txt"),
                           full.names = TRUE)) {
    texts <- c(texts, readChar(fname, file.info(fname)$size))
    labels <- c(labels, label)
  }
}

```

The data is transformed as tokens for feeding it in the neural network as a form of tensors.

```{r}

### Tokenize the data

maxlen <- 150                 # We will cut reviews after 150 words
training_samples <- 100       # We will be training on 100 samples
validation_samples <- 10000   # We will be validating on 10000 samples
max_words <- 10000            # We will only consider the top 10,000 words in the dataset

tokenizer <- text_tokenizer(num_words = max_words) %>% 
  fit_text_tokenizer(texts)

sequences <- texts_to_sequences(tokenizer, texts)
word_index = tokenizer$word_index

cat("Found", length(word_index), "unique tokens.\n")
data <- pad_sequences(sequences, maxlen = maxlen)
y_data <- as.array(labels)
labels <- as.array(labels)
cat("Shape of data tensor:", dim(data), "\n")
cat('Shape of label tensor:', dim(labels), "\n")

```

The data is partitioned in the train and validation.

```{r}

set.seed(123)
indices <- sample(1:nrow(data))
training_indices <- indices[1:training_samples]
validation_indices <- indices[(training_samples + 1): 
                                (training_samples + validation_samples)]

x_train <- data[training_indices,]
y_train <- labels[training_indices]

x_val <- data[validation_indices,]
y_val <- labels[validation_indices]

```

#Pretrain Network

GLove algorithm is used for the Pretraining network and below the preprocessing of the network is done. 

```{r}

glove_dir = 'C:/Users/12342/Documents/AdvancedML/Assignment/glove.6B'
lines <- readLines(file.path(glove_dir, "glove.6B.100d.txt"))

embeddings_index <- new.env(hash = TRUE, parent = emptyenv())
for (i in 1:length(lines)) {
  line <- lines[[i]]
  values <- strsplit(line, " ")[[1]]
  word <- values[[1]]
  embeddings_index[[word]] <- as.double(values[-1])
}

cat("Found", length(embeddings_index), "word vectors.\n")
```

Building an embedding matrix that can be loaded into an embedding layer.

```{r}
embedding_dim <- 100
embedding_matrix <- array(0, c(max_words, embedding_dim))

for (word in names(word_index)) {
  index <- word_index[[word]]
  if (index < max_words) {
    embedding_vector <- embeddings_index[[word]]
    if (!is.null(embedding_vector))
      # Words not found in the embedding index will be all zeros.
      embedding_matrix[index+1,] <- embedding_vector
  }
}
```

Neural Network model is build.

```{r}
model <- keras_model_sequential() %>% 
  layer_embedding(input_dim = max_words, output_dim = embedding_dim, 
                  input_length = maxlen) %>% 
  layer_flatten() %>% 
  layer_dense(units = 32, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")

summary(model)
```

Load the pretrained GloVe embeddings in the model

```{r}

get_layer(model, index = 1) %>% 
  set_weights(list(embedding_matrix)) %>% 
  freeze_weights()

model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("acc")
)
```

Evaluating on train model

```{r}
history <- model %>% fit(
  x_train, y_train,
  epochs = 20,
  batch_size = 32,
  validation_data = list(x_val, y_val)
)
plot(history)
```

Evaluating on test data

```{r}
model %>% fit(
  x_train,
  y_train,
  epochs = 2,
  batch_size = 32)
result <- model %>%  evaluate(data,y_data)
result # Test Acuuracy of the model is 51%
```

#Embedding layer

Using an embedding layer and classifier on the IMDB data

```{r}

model <- keras_model_sequential() %>% 
  layer_embedding(input_dim = max_words, output_dim = embedding_dim, 
                  input_length = maxlen) %>% 
  layer_flatten() %>% 
  layer_dense(units = 32, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")

model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("acc")
)
```

```{r}
history1 <- model %>% fit(
  x_train, y_train,
  epochs = 20,
  batch_size = 32,
  validation_data = list(x_val, y_val)
)

plot(history1)
```

Model execution with hypertunned parameters on the test data.

```{r}
model %>% fit(
  x_train,
  y_train,
  epochs = 2,
  batch_size = 32)
result1 <- model %>%  evaluate(data,y_data)
result1
```

#Few sets of Training samples always tend to overfitting. 
#Performance of a model depends on the number of samples we take, if we choose to have less number of samples it rely on 100 samples. 
#with less number of samples, pretrained layer and embedding layer has the same model accuracy.

